geochemistry

### 2025-01-27 geochemistry.qmd
Chapter structure (initial)
PXrf data analysis (initial tables).

@@

@@

pxrf_dataset <- pxrf_raw %>%
  select("Sample_ID", "pXRF_ID", "Formation/Type", "Application", "Method",
         "ElapsedTime",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2",
         "V", "Cr", "MnO", "Fe2O3", "Co", "Ni", "Cu", "Zn", "As", "Se", "Rb",
         "Sr", "Y", "Zr", "Nb", "Mo", "Rh", "Ag", "Cd", "Sn", "Sb", "Ba", "La",
         "Ce", "Hf", "Ta", "W", "Pt", "Au", "Hg", "Tl", "Pb", "Bi", "Th", "U",
         "Cal Check") %>% 
  mutate("MgO_detect" = case_when(
    MgO == 0 | MgO == "< LOD" ~ "LOD",
    TRUE ~ "Value"))

pxrf_analysis <- pxrf_dataset %>% 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb") # Removed elements with only 0 or <LOD values in all samples

>>>>

library(janitor)
library(tidyr)

@@

# X-ray diffraction (XRD)

# Scanning electron microscopy and energy dispersive X-ray spectroscopy (SEM-EDS)

# Portable X-ray fluorescence (pXRF)

@@

pxrf_dataset <- pxrf_raw |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type", "Application", "Method",
         "ElapsedTime",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2",
         "V", "Cr", "MnO", "Fe2O3", "Co", "Ni", "Cu", "Zn", "As", "Se", "Rb",
         "Sr", "Y", "Zr", "Nb", "Mo", "Rh", "Ag", "Cd", "Sn", "Sb", "Ba", "La",
         "Ce", "Hf", "Ta", "W", "Pt", "Au", "Hg", "Tl", "Pb", "Bi", "Th", "U",
         "Cal Check") |> 
  mutate("MgO_detect" = case_when(
    MgO == 0 | MgO == "< LOD" ~ "LOD",
    TRUE ~ "Value"))

pxrf_analysis <- pxrf_dataset |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb") # Removed elements with only 0 or <LOD values in all samples

```{r}
#| label: tbl-occurrences
#| tbl-cap: A table.

# Example dataset
data <- data.frame(
  Group = c("A", "A", "B", "B", "A", "B", "C", "C", "C", "A"),
  Category1 = c("X", "Y", "X", "Y", "X", "Y", "X", "X", "Y", "Y"),
  Category2 = c("P", "P", "Q", "Q", "P", "P", "Q", "P", "Q", "P")
)

# Combine the multiple columns into a single table
combined_table <- pxrf_analysis |> 
  mutate(across(
    .cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "V", "MnO", "Fe2O3", "Cu", "As", "Rb",
         "Sr", "Y", "Zr", "Mo", "Ag", "Ba", "Ce", "W", "Au", "Pb"),
    .fns = ~ case_when(
      . == 0 | . == "< LOD" ~ "No",
      . > 0 ~ "Yes",
      TRUE ~ as.character(.))
    )) |> 
  mutate(`Chert type` = case_when(
    `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "TYPE 2",
    `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "TYPE 6",
    TRUE ~ `Formation/Type`))

tbl_MgO <- combined_table |> 
  tabyl(`Chert type`, MgO) |>
  adorn_percentages("row") |> 
  adorn_pct_formatting(digits = 1) |> 
  select("Chert type", Yes) |> 
  rename(MgO = Yes)
# Al2O3
tbl_Al2O3 <- combined_table |> 
  tabyl(`Chert type`, Al2O3) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>
  select("Chert type", Yes) |> 
  rename(Al2O3 = Yes)
# SiO2
tbl_SiO2 <- combined_table |> 
  tabyl(`Chert type`, SiO2) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>
  select("Chert type", Yes) |> 
  rename(SiO2 = Yes)
# P2O5
tbl_P2O5 <- combined_table |> 
  tabyl(`Chert type`, P2O5) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(P2O5 = Yes)
# S
tbl_S <- combined_table |> 
  tabyl(`Chert type`, S) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(S = Yes)
# Cl
tbl_Cl <- combined_table |> 
  tabyl(`Chert type`, Cl) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Cl = Yes)
# K2O has 100% rate
tbl_K2O <- combined_table |> 
  tabyl(`Chert type`, K2O) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(K2O = Yes)
# CaO
tbl_CaO <- combined_table |> 
  tabyl(`Chert type`, CaO) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(CaO = Yes)
# TiO2
tbl_TiO2 <- combined_table |> 
  tabyl(`Chert type`, TiO2) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(TiO2 = Yes)
# V
tbl_V <- combined_table |> 
  tabyl(`Chert type`, V) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(V = Yes)
# MnO
tbl_MnO <- combined_table |> 
  tabyl(`Chert type`, MnO) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(MnO = Yes)
# Fe2O3 has 100% rate
tbl_Fe2O3 <- combined_table |> 
  tabyl(`Chert type`, Fe2O3) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Fe2O3 = Yes)
# Cu
tbl_Cu <- combined_table |> 
  tabyl(`Chert type`, Cu) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Cu = Yes)
# As
tbl_As <- combined_table |> 
  tabyl(`Chert type`, As) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(As = Yes)
# Rb
tbl_Rb <- combined_table |> 
  tabyl(`Chert type`, Rb) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Rb = Yes)
# Sr
tbl_Sr <- combined_table |> 
  tabyl(`Chert type`, Sr) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Sr = Yes)
# Y
tbl_Y <- combined_table |> 
  tabyl(`Chert type`, Y) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Y = Yes)
# Zr
tbl_Zr <- combined_table |> 
  tabyl(`Chert type`, Zr) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Zr = Yes)
# Mo
tbl_Mo <- combined_table |> 
  tabyl(`Chert type`, Mo) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Mo = Yes)
# Ag
tbl_Ag <- combined_table |> 
  tabyl(`Chert type`, Ag) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Ag = Yes)
# Ba
tbl_Ba <- combined_table |> 
  tabyl(`Chert type`, Ba) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Ba = Yes)
# Ce
tbl_Ce <- combined_table |> 
  tabyl(`Chert type`, Ce) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Ce = Yes)
# W
tbl_W <- combined_table |> 
  tabyl(`Chert type`, W) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(W = Yes)
# Au
tbl_Au <- combined_table |> 
  tabyl(`Chert type`, Au) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Au = Yes)
# Pb
tbl_Pb <- combined_table |> 
  tabyl(`Chert type`, Pb) |>
  adorn_percentages("row") |>                   
  adorn_pct_formatting(digits = 1) |>

  select("Chert type", Yes) |> 
  rename(Pb = Yes)

combined_tbl <- tbl_MgO |> 
  full_join(tbl_Al2O3, by = "Chert type") |> 
  full_join(tbl_SiO2, by = "Chert type") |>
  full_join(tbl_P2O5, by = "Chert type") |> 
  full_join(tbl_S, by = "Chert type") |> 
  full_join(tbl_Cl, by = "Chert type") |> 
  full_join(tbl_K2O, by = "Chert type") |> 
  full_join(tbl_CaO, by = "Chert type") |> 
  full_join(tbl_TiO2, by = "Chert type") |> 
  full_join(tbl_V, by = "Chert type") |> 
  full_join(tbl_MnO, by = "Chert type") |> 
  full_join(tbl_Fe2O3, by = "Chert type") |> 
  full_join(tbl_Cu, by = "Chert type") |> 
  full_join(tbl_As, by = "Chert type") |> 
  full_join(tbl_Rb, by = "Chert type") |> 
  full_join(tbl_Sr, by = "Chert type") |> 
  full_join(tbl_Y, by = "Chert type") |> 
  full_join(tbl_Zr, by = "Chert type") |> 
  full_join(tbl_Mo, by = "Chert type") |> 
  full_join(tbl_Ag, by = "Chert type") |> 
  full_join(tbl_Ba, by = "Chert type") |> 
  full_join(tbl_Ce, by = "Chert type") |> 
  full_join(tbl_W, by = "Chert type") |> 
  full_join(tbl_Au, by = "Chert type") |> 
  full_join(tbl_Pb, by = "Chert type")

library(gt)

combined_table |> 
  select(`Formation/Type`, MgO, Al2O3) |> 
  t

# Print the resulting table
print(combined_table)
